---
title: "5210assignment2"
output: 
  pdf_document: 
date: "2023-11-27"
---

# Q1:

```{r import data}
data1 <- read.table(header = TRUE, text="
patient age acid x-ray size grade y
1 66 0.48 0 0 0 0
2 68 0.56 0 0 0 0
3 66 0.50 0 0 0 0
4 56 0.52 0 0 0 0
5 58 0.50 0 0 0 0
6 60 0.49 0 0 0 0
7 65 0.46 1 0 0 0
8 60 0.62 1 0 0 0
9 50 0.56 0 0 1 1
10 49 0.55 1 0 0 0
11 61 0.62 0 0 0 0
12 58 0.71 0 0 0 0
13 51 0.65 0 0 0 0
14 67 0.67 1 0 1 1
15 67 0.47 0 0 1 0
16 51 0.49 0 0 0 0
17 56 0.50 0 0 1 0
18 60 0.78 0 0 0 0
19 52 0.83 0 0 0 0
20 56 0.98 0 0 0 0
21 67 0.52 0 0 0 0
22 63 0.75 0 0 0 0
23 59 0.99 0 0 1 1
24 64 1.87 0 0 0 0
25 61 1.36 1 0 0 1
26 56 0.82 0 0 0 1
27 64 0.40 0 1 1 0
28 61 0.50 0 1 0 0
29 64 0.50 0 1 1 0
30 63 0.40 0 1 0 0
31 52 0.55 0 1 1 0
32 66 0.59 0 1 1 0
33 58 0.48 1 1 0 1
34 57 0.51 1 1 1 1
35 65 0.49 0 1 0 1
36 65 0.48 0 1 1 0
37 59 0.63 1 1 1 0
38 61 1.02 0 1 0 0
39 53 0.76 0 1 0 0
40 67 0.95 0 1 0 0
41 53 0.66 0 1 1 0
42 65 0.84 1 1 1 1
43 50 0.81 1 1 1 1
44 60 0.76 1 1 1 1
45 45 0.70 0 1 1 1
46 56 0.78 1 1 1 1
47 46 0.70 0 1 0 1
48 67 0.67 0 1 0 1
49 63 0.82 0 1 0 1
50 57 0.67 0 1 1 1
51 51 0.72 1 1 0 1
52 64 0.89 1 1 0 1
53 68 1.26 1 1 1 1
")

```

```{r delete useless variable}
#delete useless variable
data1$patient <- NULL
```

(a)Show the results from each of the 4 steps in table format leading to the selection of your final model. Include the variable tested, the deviance and df.

## Step1:

```{r Step 1}
#Fit single-variable model
variables1 <- c("age", "acid", "x.ray", "size", "grade")
single_var_models <- lapply(variables1, function(var) {
                     glm(formula = paste("y ~", var), data = data1, family = binomial)
})
print(single_var_models)
```

```{r step1 Compare deviance}
# Compare deviance of each model
null_model <- glm(y ~ 1, data = data1, family = binomial)
single_var_deviance <- sapply(single_var_models, function(model) {
                       deviance(model) - deviance(null_model)
})
print(single_var_deviance)
```

We can perform Square Root Transformation, Square Transformation and Log transformatin on variables(age and acid).

```{r step1 transformation & uni-variable model, echo=TRUE, message=FALSE}
# SRT
data1$sqrt_age <- sqrt(data1$age)
data1$sqrt_acid <- sqrt(data1$acid)
# ST
data1$sq_age <- data1$age^2
data1$sq_acid <- data1$acid^2
#Log
data1$log_age <- log(data1$age )
data1$log_acid <- log(data1$acid )


#Fit single variable model
variables2 <- c("sqrt_age", "sqrt_acid", "sq_age", "sq_acid", "log_age", "log_acid")
single_var_models2 <- lapply(variables2, function(var) {
                     glm(formula = paste("y ~", var), data = data1, family = binomial)
})
print(single_var_models2)

```

```{r compare deviance}
# Compare deviance of each model
null_model <- glm(y ~ 1, data = data1, family = binomial)
single_var_deviance2 <- sapply(single_var_models2, function(model) {
                       deviance(model) - deviance(null_model)
})
print(single_var_deviance2)
print(null_model)
```

## Step2:

```{r step2}
variables3 <- c("log_acid", "x.ray", "size", "grade")
formula <- as.formula(paste("y ~", paste(variables3, collapse = " + ")))
multi_var_model1 <- glm(formula, family = binomial(link = "logit"), data = data1)

print(multi_var_model1)
```

```{r reduced_model}
for (variable in variables3) {
   reduced_formula <- as.formula(paste("y ~", paste(setdiff(variables3, variable),
                                                    collapse = " + ")))

reduced_model <- glm(reduced_formula, family = binomial(link = "logit"), data = data1)

print(reduced_model)

}
```

```{r reduced model1}
variables4 <- c("x.ray", "size", "log_acid")
formula1 <- as.formula(paste("y ~", paste(variables4, collapse = " + ")))
multi_var_model2 <- glm(formula1, family = binomial(link = "logit"), data = data1)

print(multi_var_model2)
for (variable in variables4) {
   reduced_formula <- as.formula(paste("y ~", paste(setdiff(variables4, variable),
                                                    collapse = " + ")))

reduced_model1 <- glm(reduced_formula, family = binomial(link = "logit"), data = data1)

print(reduced_model1)
}
```

## step3:

```{r}
inter_model <- glm( y ~ log_acid + x.ray + size ,
                     family = binomial, data = data1)
print(inter_model)
```

```{r interaction 3}
inter_model1 <- glm( y ~ log_acid + x.ray + size + 
                       log_acid * x.ray * size,
                     family = binomial, data = data1)
print(inter_model1)
```

```{r}
inter_model2 <- glm( y ~ log_acid + x.ray + size + 
    x.ray * size + x.ray * log_acid + size * log_acid
    , family = binomial, data = data1)
print(inter_model2)

inter_model2_1 <- glm( y ~ log_acid + x.ray + size + 
    x.ray * size 
    , family = binomial, data = data1)
print(inter_model2_1)

inter_model2_2 <- glm( y ~ log_acid + x.ray + size + 
     x.ray * log_acid 
    , family = binomial, data = data1)
print(inter_model2_2)

inter_model2_3 <- glm( y ~ log_acid + x.ray + size + 
     size * log_acid
    , family = binomial, data = data1)
print(inter_model2_3)

inter_model2_4 <- glm( y ~ log_acid + x.ray + size + 
     age
    , family = binomial, data = data1)
print(inter_model2_4)
```

```{r}
inter_model3_1 <- glm( y ~ log_acid + x.ray + size + 
     x.ray * log_acid + size * log_acid
    , family = binomial, data = data1)
print(inter_model3_1)

inter_model3_2 <- glm( y ~ log_acid + x.ray + size + 
    x.ray * size + size * log_acid
    , family = binomial, data = data1)
print(inter_model3_2)

inter_model3_3 <- glm( y ~ log_acid + x.ray + size + 
    x.ray * size + x.ray * log_acid 
    , family = binomial, data = data1)
print(inter_model3_3)
```

# Q4

a)  Write a SAS or R program that a) inputs the data into SAS or R and b) fits a cumulative probability model where the only predictor is political party

```{r input data}
# Library package
library(MASS)
library(dplyr)
library(tidyr)

# Create data frame
data4 <- data.frame(
  Gender = rep(c("Female", "Male"), each = 2),
  Party = rep(c("Democratic", "Republican"), each = 1),
  VeryLiberal = c(44, 18, 36, 12),
  SlightlyLiberal = c(47, 28, 34, 18),
  Moderate = c(118, 86, 53, 62),
  SlightlyConservative = c(23, 39, 18, 45),
  VeryConservative = c(32, 48, 23, 51)
)


```

```{r}
long_data4 <- data4 %>%
  pivot_longer(cols = VeryLiberal:VeryConservative,
               names_to = "Ideology",
               values_to = "Count") %>%
  mutate(PartyInd = as.numeric(Party == "Democratic"),
         Ideology = factor(Ideology, levels = c("VeryLiberal", "SlightlyLiberal", "Moderate", "SlightlyConservative", "VeryConservative"), ordered = TRUE))

# Fit cumulative logit model
model <- polr(Ideology ~ PartyInd, weights = Count, data = long_data4, Hess=TRUE)
summary(model)
```

# 
